---
title: "sex ratio script"
output: html_document
date: "2023-06-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load in phytoplankton and water temp data
```{r} 
phyto_1m <- read.csv("BC_Phytoplankton_New-1m_Wholewater_L_Quant.csv")
phyto_1m
waterconditions <- read.csv("BC_Phytoplankton_New-Conditions.csv")
waterconditions
```

#load in raw sex ratio data from Excel
```{r}
Pacific_Oyster_Sex_Ratio <- read.csv("Pacific_Oyster_Sex_Ratio.csv")
Pacific_Oyster_Sex_Ratio

#the real stuff
POSR_REAL <- read.csv("Pacific_Oyster_Sex_Ratio_REAL.csv")
POSR_REAL
```

#update values in Environment; totalfemale, totalherm, femalegonads, percents female watertempsfake, total plankton
```{r}
totalfemale = Pacific_Oyster_Sex_Ratio$Female
totalherm = Pacific_Oyster_Sex_Ratio$Hermaphrodite
femalegonads = totalfemale + totalherm
percentsfemale = femalegonads/Pacific_Oyster_Sex_Ratio$`Total Oysters Collected`
totalplankton = phyto_1m$Diatom_Sum[59:66]
#note: will need to change to [85:93] but excluding 86
totalplankton
watertempsfake = waterconditions$Water_Temp_1m_C[79:86]
#note: will need to change to [105:115] excluding 106, 110, and 111
watertempsfake

#the real stuff
totalfemale_REAL = POSR_REAL$Female
totalherm_REAL = POSR_REAL$Hermaphrodite
femalegonads_REAL = totalfemale_REAL + totalherm_REAL
percentsfemale_REAL = femalegonads_REAL/POSR_REAL$Total.Oysters.Collected
totaldiatom <- phyto_1m[c(84,86,87,88,89,90), "Diatom_Sum"]
totaldiatom
watertemps <- waterconditions[c(104,106,107,108,111,112), "Water_Temp_1m_C"]
watertemps
#make sure that row names are adjusted as new data comes in
```

#add new columns to data sheet; Total Female Gonads, Percentage Female Gonads, Water Temps at 1m, ID_yyyymmdd
```{r}
Pacific_Oyster_Sex_Ratio$Total_Female_Gonads <- femalegonads
percentsfemale = femalegonads/Pacific_Oyster_Sex_Ratio$Total.Oysters.Collected
Pacific_Oyster_Sex_Ratio$Percentage_Female_Gonads <- percentsfemale
Pacific_Oyster_Sex_Ratio$Water_Temps_at_1m <- watertempsfake
Pacific_Oyster_Sex_Ratio$Total_Plankton <- totalplankton
datesID <- c("20230606", "20230613", "20230620", "20230626", "20230705", "20230711", "20230718", "20230725")
Pacific_Oyster_Sex_Ratio$ID_yyyymmdd <- datesID
juliandates <- c(23157, 23164, 23171, 23177, 23186, 23192, 23199, 23206)
Pacific_Oyster_Sex_Ratio$Julian_Date <- juliandates
Pacific_Oyster_Sex_Ratio

#the real stuff
POSR_REAL$Total_Female_Gonads <- femalegonads_REAL
POSR_REAL$Percentage_Female_Gonads <- percentsfemale_REAL
POSR_REAL$Water_Temp_at_1m <- watertemps
POSR_REAL$Total_Diatom <- totaldiatom
datesID <- c("20230606", "20230613", "20230620", "20230626", "20230705", "20230711")
POSR_REAL$ID_yyyymmdd <- datesID
juliandates <- c(23157, 23164, 23171, 23177, 23186, 23192)
POSR_REAL$Julian_Date <- juliandates
POSR_REAL
```

#rearrange column order
```{r}
library(data.table)
setcolorder(Pacific_Oyster_Sex_Ratio, c("Date.Collected", "ID_yyyymmdd", "Julian_Date", "Total.Oysters.Collected", "Male", "Female", "Hermaphrodite", "Total_Female_Gonads", "Percentage_Female_Gonads", "Water_Temps_at_1m", "Total_Plankton"))
Pacific_Oyster_Sex_Ratio

#the real stuff
library(data.table)
setcolorder(POSR_REAL, c("Date.Collected", "ID_yyyymmdd", "Julian_Date", "Total.Oysters.Collected", "Male", "Female", "Hermaphrodite", "Total_Female_Gonads", "Percentage_Female_Gonads", "Water_Temp_at_1m", "Total_Diatom"))
POSR_REAL
```

#plot sex ratio over time
```{r}
Pacific_Oyster_Sex_Ratio$ID_yyyymmdd <- as.numeric(Pacific_Oyster_Sex_Ratio$ID_yyyymmdd)
Pacific_Oyster_Sex_Ratio$Dates <- as.Date(Pacific_Oyster_Sex_Ratio$Date.Collected, format="%m/%d/%y")
plot(Pacific_Oyster_Sex_Ratio$Dates, Pacific_Oyster_Sex_Ratio$Percentage_Female_Gonads, ylab="Percentage of Female Gonads", xlab= "Time")

#the real stuff
POSR_REAL$ID_yyyymmdd <- as.numeric(POSR_REAL$ID_yyyymmdd)
POSR_REAL$Dates <- as.Date(POSR_REAL$Date.Collected,
                           format = "%m/%d/%y")
POSR_REAL
plot(POSR_REAL$Dates,
     POSR_REAL$Percentage_Female_Gonads,
     ylab = "Percentage of Oysters with Ova Present",
     xlab = "Time")
# instead of "Percentage of Female Gonads", "Ova Present"?

#ggplot time
install.packages("ggplot2")
library(ggplot2)
graph1 <- qplot(Dates, Percentage_Female_Gonads, data = POSR_REAL, xlab = "Date Collected", ylab="Percentage of Oysters with Ova Present")
ptcolor <- 'grey20'
graph1 +
  ggtitle("Percentage of Oysters Collected with Ova Present Throughout Study") +
  theme(plot.title=element_text(size=14, lineheight=0.8, color=ptcolor, hjust=0.5),
    axis.title.x=element_text(color=ptcolor),
    axis.title.y=element_text(color=ptcolor))

#create a common theme to reuse throughout plots
common_theme <- function() {
  ptcolor <- 'grey20' # plot text color
  theme(
    plot.title=element_text(size=14, lineheight=0.8, color=ptcolor, hjust=0.5),
    axis.title.x=element_text(color=ptcolor),
    axis.title.y=element_text(color=ptcolor))
}

graph1 +
  ggtitle("Percentage of Oysters Collected with Ova Present Throughout Study") +
  theme_bw()+
  theme(plot.title=element_text(size=14, lineheight=0.8, color=ptcolor, hjust=0.5))

```

#plot sex ratio over water temps
```{r}
plot(POSR_REAL$Water_Temp_at_1m, POSR_REAL$Percentage_Female_Gonads, ylab="Percentage of Oysters with Ova Present", xlab="Water Temperature at 1m")

graph2 <- qplot(Water_Temp_at_1m, 
      Percentage_Female_Gonads, 
      data=POSR_REAL, 
      xlab = "Water Temperature at 1m (degrees C)", 
      ylab="Oysters with Ova Present (% Collected)")

graph2+
  ggtitle("Oysters Collected with Ova Present and Water Temperature")+
  theme_bw()+
  theme(plot.title=element_text(size=14, lineheight=0.8, color=ptcolor, hjust=0.5))
```

#plot sex ratio over phytoplankton abundance
```{r}
plot(POSR_REAL$Total_Diatom, POSR_REAL$Percentage_Female_Gonads, ylab="Percentage of Female Gonads", xlab="Diatom Abundance at 1m")

graph3 <- qplot(Total_Diatom, 
      Percentage_Female_Gonads, 
      data=POSR_REAL, 
      xlab="Diatom Abundance at 1m", 
      ylab="Percentage of Oysters with Ova Present")

graph3+
  ggtitle("Oysters Collected with Ova Present and Diatom Abundance")+
  theme_bw()+
  theme(plot.title=element_text(size=14, lineheight=0.8, color=ptcolor, hjust=0.5))
```
# note: water temps and total plankton data are from corresponding dates from 2022 (not exact dates, but same week)

#create input vectors for barplot and update barplot values
```{r}
colors = c("lightblue", "pink", "yellow")
regions <- c("Male", "Female", "Hermaphrodite")
barplotdates = c("6/6", "6/13", "6/20", "6/26", "7/5", "7/11")
barplotvalues <- matrix(data = c(POSR_REAL$Male, POSR_REAL$Female, POSR_REAL$Hermaphrodite), nrow = 3, ncol = 6, byrow = TRUE)

#the real stuff
```

#plot sex ratio bar plot
```{r}
barplot(barplotvalues, main= "Sex Ratio Change During Summer Months", xlab = "Date Collected", ylab="Number of Oysters", col=colors, names.arg=barplotdates)
legend("topleft", regions, cex = 0.4, fill = colors)
```
```{r}
#ggplot2 stacked bar plot
regions <- c("Male", "Female", "Hermaphrodite")
barplotdates = c("6/6", "6/13", "6/20", "6/26", "7/5", "7/11")
barplotdata <- data.frame(barplotvalues+barplotdates, row.names=regions)
barplotdata
ggplot(data=barplotdata, aes(fill=c(Male, Female, Hermaphrodite)))+
  geom_bar(stat='identity', position=position_stack())
```


```{r}
#linear model time
#I want to look at the relationship between water temperature and percentage of female gonads; and diatom abundance and female gonads; time and percentage of female gonads
#water temperature
#dependent variable as predicted by independent: gonads~watertemp
#percentage of female gonads vs. water temp at 1 m
m1.binomial <- glm(Percentage_Female_Gonads~Water_Temp_at_1m, 
                   family = "binomial",
                   data = POSR_REAL)
summary(m1.binomial)

#percentage of female gonads vs. total diatom abundance at 1m
m2.binomial <- glm(Percentage_Female_Gonads~Water_Temp_at_1m+Total_Diatom, 
                   family = "binomial",
                   data = POSR_REAL)
summary(m2.binomial)

#AIC: does adding diatom abundance make the data a better fit? or make the model too complicated?
AIC(m1.binomial, m2.binomial)

#percentage of female gonads vs. time
m3.binomial <- glm(Percentage_Female_Gonads~Dates,
                   family = "binomial",
                   data = POSR_REAL)
summary(m3.binomial)

m4.binomial <- glm(Percentage_Female_Gonads~Julian_Date,
                   family = "binomial",
                   data = POSR_REAL)
summary(m4.binomial)
```

