---
title: "sex ratio script"
output: html_document
date: "2023-06-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load in phytoplankton and water temp data
```{r} 
phyto_1m <- read.csv("BC_Phytoplankton_New-1m_Wholewater_L_Quant(1).csv")
phyto_1m
waterconditions <- read.csv("BC_Phytoplankton_New-Conditions.csv")
waterconditions
```

#load in raw sex ratio data from Excel
```{r}
Pacific_Oyster_Sex_Ratio <- read.csv("Pacific_Oyster_Sex_Ratio.csv")
Pacific_Oyster_Sex_Ratio
```

#update values in Environment; totalfemale, totalherm, femalegonads, percents female watertempsfake, total plankton
```{r}
totalfemale = Pacific_Oyster_Sex_Ratio$Female
totalherm = Pacific_Oyster_Sex_Ratio$Hermaphrodite
femalegonads = totalfemale + totalherm
percentsfemale = femalegonads/Pacific_Oyster_Sex_Ratio$`Total Oysters Collected`
totalplankton = phyto_1m$Diatom_Sum[59:66]
#note: will need to change to [85:93] but excluding 86
totalplankton
watertempsfake = waterconditions$Water_Temp_1m_C[79:86]
#note: will need to change to [105:115] excluding 106, 110, and 111
watertempsfake
```

#add new columns to data sheet; Total Female Gonads, Percentage Female Gonads, Water Temps at 1m, ID_yyyymmdd
```{r}
Pacific_Oyster_Sex_Ratio$Total_Female_Gonads <- femalegonads
percentsfemale = femalegonads/Pacific_Oyster_Sex_Ratio$Total.Oysters.Collected
Pacific_Oyster_Sex_Ratio$Percentage_Female_Gonads <- percentsfemale
Pacific_Oyster_Sex_Ratio$Water_Temps_at_1m <- watertempsfake
Pacific_Oyster_Sex_Ratio$Total_Plankton <- totalplankton
datesID <- c("20230606", "20230613", "20230620", "20230626", "20230705", "20230711", "20230718", "20230725")
Pacific_Oyster_Sex_Ratio$ID_yyyymmdd <- datesID
Pacific_Oyster_Sex_Ratio
```

#rearrange column order
```{r}
library(data.table)
setcolorder(Pacific_Oyster_Sex_Ratio, c("Date.Collected", "ID_yyyymmdd", "Total.Oysters.Collected", "Male", "Female", "Hermaphrodite", "Total_Female_Gonads", "Percentage_Female_Gonads", "Water_Temps_at_1m", "Total_Plankton"))
Pacific_Oyster_Sex_Ratio
```

#plot sex ratio over time
```{r}
plot(Pacific_Oyster_Sex_Ratio$``, Pacific_Oyster_Sex_Ratio$Percentage_Female_Gonads, ylab="Percentage of Female Gonads", xlab= "Time")
#how do I fix the x values so they go day by day but are still integers...
```

#plot sex ratio over water temps
```{r}
plot(Pacific_Oyster_Sex_Ratio$Water_Temps_at_1m, Pacific_Oyster_Sex_Ratio$Percentage_Female_Gonads, ylab="Percentage of Female Gonads", xlab="Water Temperature at 1m")
```

#plot sex ratio over phytoplankton abundance
```{r}
plot(Pacific_Oyster_Sex_Ratio$Total_Plankton, Pacific_Oyster_Sex_Ratio$Percentage_Female_Gonads, ylab="Percentage of Female Gonads", xlab="Diatom Abundance at 1m")
```
# note: water temps and total plankton data are from corresponding dates from 2022 (not exact dates, but same week)

#create input vectors for barplot and update barplot values
```{r}
colors = c("lightblue", "pink", "yellow")
regions <- c("Male", "Female", "Hermaphrodite")
barplotdates = c("6/6", "6/13", "6/20", "6/26", "7/5", "7/11", "7/18", "7/25")
barplotvalues <- matrix(data = c(Pacific_Oyster_Sex_Ratio$Male, Pacific_Oyster_Sex_Ratio$Female, Pacific_Oyster_Sex_Ratio$Hermaphrodite), nrow = 3, ncol = 8, byrow = TRUE)
```

#plot sex ratio bar plot
```{r}
barplot(barplotvalues, main= "Sex Ratio Change During Summer Months", xlab = "Date Collected", ylab="Sex Ratio", col=colors, names.arg=barplotdates)
legend("topleft", regions, cex = 0.4, fill = colors)
```


